## Chenxi Liu
## Assightment 7

```{r, echo=FALSE,warning=FALSE,include=FALSE}

library(dplyr, warn.conflicts = F, quietly = T)
library(tidyr, warn.conflicts = F, quietly = T)
library(readxl, warn.conflicts = F, quietly = T)
library(janitor, warn.conflicts = F, quietly = T)
library(tidyverse, warn.conflicts = F, quietly = T)
library(ggplot2, warn.conflicts = F, quietly = T)
library(sf, warn.conflicts = F, quietly = T)
library(spData, warn.conflicts = F, quietly = T)
library(maps, warn.conflicts = F, quietly = T)
library(viridis, warn.conflicts = F, quietly = T)
library(ggpubr, warn.conflicts = F, quietly = T)
library(cowplot, warn.conflicts = F, quietly = T)
library(stringr, warn.conflicts = F, quietly = T)
library(reticulate, warn.conflicts = F, quietly = T)
```


```{r, echo=FALSE,warning=FALSE,include=FALSE}
py_install('pandas')
py_install('numpy')
py_install('matplotlib')
py_install('plotly')
py_install('xlrd')
py_install('networkx')
py_install('openpyxl')
```


Analysis note: 

I choose California and Alaska to be my two extra states while they are a lot of people moving away from Texas, California, and New York. People who live in Alaska seems to not move that much. (Maybe due to the small population based on other three states) The arrow indicates the direction where people are moving. Four different colors indicates four origins that we choose. The thickness of the arrow indicates the number of people. 



```{python, echo=FALSE,warning=FALSE,include = FALSE}
import openpyxl
import pandas as pd 
import numpy as np
import networkx as nx
import matplotlib.pyplot as plt
import random 

import os
os.environ['QT_QPA_PLATFORM_PLUGIN_PATH'] = 'C:/Users/liuch/AppData/Local/r-miniconda/envs/r-reticulate/Library/plugins/platforms'
file = 'State_to_State_Migrations_Table_2018.xls'
data = pd.ExcelFile(file)
df = data.parse('Table')

#### get the starting point order 
state = []
for i in df.iloc[5]:
    if i is not np.nan and i != 'Total':
        state.append(i)
        
state = state[:len(state)-2]

# drop the rows that is not relevant

for i in range(10):
    df = df.drop(i)

# reset index after dropping
    
df.reset_index(drop=True, inplace=True)

# create a list to drop 

empty_row_index = []

# drop empty rows 
for i in range(len(df)-1):
    if df.iloc[i][0] is np.nan:
        empty_row_index.append(i)

for i in empty_row_index:
    df = df.drop(i)
    
# reset index after dropping 
df.reset_index(drop=True, inplace=True)    

# drop the rows that is not relevant
for i in range(53,len(df)):
    df = df.drop(i)

# reset index after dropping 
df.reset_index(drop=True, inplace=True)    

# drop the row with header 
header_index = []

for i in range(len(df)-1):
    if df.iloc[i][0] not in state:
        header_index.append(i)


for i in header_index:
    df = df.drop(i)
    
# reset index after dropping 
df.reset_index(drop=True, inplace=True)      



delete_col = []

for i in range(1,9):
    delete_col.append(df.columns[i])

for i in df.columns:
    if df[i][0] in state and i is not df.columns[0]:
        delete_col.append(i)
for i in delete_col:
    df = df.drop(i, axis=1)

second_delete_col = []

for i in range(len(df.columns)):
    if i is not 0 and i%2 == 0:
        second_delete_col.append(df.columns[i])
second_delete_col.append('Unnamed: 128')
second_delete_col.append('Unnamed: 126')
second_delete_col.append('Unnamed: 122')
for i in second_delete_col:
    df = df.drop(i, axis=1)

new_columns = ['State'] + state
df.columns = new_columns



target = []
start = []
value = []

for i in range(len(df)):
    for col in df.columns:
        if df[col][i] is not np.nan and col != 'State' and col != 'Puerto Rico' and df['State'][i] != 'Puerto Rico':
            target.append(df['State'][i])
            start.append(col)
            value.append(df[col][i])


final_df = pd.DataFrame(list(zip(start, target, value)), 
               columns =['Source', 'Target', 'Count']) 
cali_df = final_df[final_df['Source'] == 'California']
texas_df = final_df[final_df['Source'] == 'Texas']
ny_df = final_df[final_df['Source'] == 'New York']
alaska_df = final_df[final_df['Source'] == 'Alaska']

cali_df = cali_df.sort_values('Count').head(10)
texas_df = texas_df.sort_values('Count').head(10)
ny_df = ny_df.sort_values('Count').head(10)
alaska_df = alaska_df.sort_values('Count').head(10)

cali_df.reset_index(drop=True, inplace=True)      
texas_df.reset_index(drop=True, inplace=True)      
ny_df.reset_index(drop=True, inplace=True)      
alaska_df.reset_index(drop=True, inplace=True)      

node_list = []
for i in range(len(cali_df)):
    if cali_df['Source'][i] not in node_list:
        node_list.append(cali_df['Source'][i])
    if cali_df['Target'][i] not in node_list:
        node_list.append(cali_df['Target'][i])
for i in range(len(texas_df)):
    if texas_df['Source'][i] not in node_list:
        node_list.append(texas_df['Source'][i])
    if texas_df['Target'][i] not in node_list:
        node_list.append(texas_df['Target'][i])
for i in range(len(ny_df)):
    if ny_df['Source'][i] not in node_list:
        node_list.append(ny_df['Source'][i])
    if ny_df['Target'][i] not in node_list:
        node_list.append(ny_df['Target'][i])
for i in range(len(alaska_df)):
    if alaska_df['Source'][i] not in node_list:
        node_list.append(alaska_df['Source'][i])
    if alaska_df['Target'][i] not in node_list:
        node_list.append(alaska_df['Target'][i])
    
G = nx.DiGraph()    
for i in node_list:
    G.add_node(i)

for i in range(len(cali_df)):
   G.add_edge(cali_df['Source'][i], cali_df['Target'][i], weight = cali_df['Count'][i], color = 'r') 
   
for i in range(len(texas_df)):
   G.add_edge(texas_df['Source'][i], texas_df['Target'][i], weight = texas_df['Count'][i], color = 'g') 
   
for i in range(len(ny_df)):
   G.add_edge(ny_df['Source'][i], ny_df['Target'][i], weight = ny_df['Count'][i], color = 'b') 
   
for i in range(len(alaska_df)):
   G.add_edge(alaska_df['Source'][i], alaska_df['Target'][i], weight = alaska_df['Count'][i], color = 'y') 

pos = nx.circular_layout(G)

edges = G.edges()
colors = []
def randomize():
    r = random.uniform(0, 1)
    g = random.uniform(0, 1)
    b = random.uniform(0, 1)
    t = (r,g,b)
    if t in colors:
        t = randomize()
    return t 
for i in range(len(G.nodes())):
    t = randomize()
    colors.append(t)
weights = [G[u][v]['weight'] for u,v in edges]
norm_weights = [float(i)*50/sum(weights) for i in weights]
edge_colors = nx.get_edge_attributes(G,'color').values()
nx.draw_networkx_nodes(G, pos, node_size = 500, node_color = colors)
nx.draw_networkx_edges(G, pos,node_size=300, arrowstyle="->", arrowsize=10,width = norm_weights, arrows = True, edge_color = edge_colors)
nx.draw_networkx_labels(G, pos, font_size=8)
```


```{python, echo=FALSE,warning=FALSE}
plt.show()
```



