
## Chenxi Liu
## Assightment 7


```{r,echo=FALSE,warning=FALSE,include=FALSE}

library(spData, warn.conflicts = F, quietly = T)
library(maps, warn.conflicts = F, quietly = T)
library(viridis, warn.conflicts = F, quietly = T)
library(ggpubr, warn.conflicts = F, quietly = T)
library(cowplot, warn.conflicts = F, quietly = T)
library(stringr, warn.conflicts = F, quietly = T)
library(reticulate, warn.conflicts = F, quietly = T)
library(dplyr, warn.conflicts = F, quietly = T)
library(tidyr, warn.conflicts = F, quietly = T)
library(readxl, warn.conflicts = F, quietly = T)
library(janitor, warn.conflicts = F, quietly = T)
library(tidyverse, warn.conflicts = F, quietly = T)
library(ggplot2, warn.conflicts = F, quietly = T)
library(sf, warn.conflicts = F, quietly = T)
```



```{r,echo=FALSE,warning=FALSE,include=FALSE}
py_install('matplotlib')
py_install('plotly')
py_install('xlrd')
py_install('pandas')
py_install('numpy')
```

```{python,echo=FALSE,warning=FALSE,include=FALSE}
import numpy as np
import matplotlib.pyplot as plt
import plotly.express as px
import plotly.offline as py
import pandas as pd
import xlrd
```


```{python,echo=FALSE,warning=FALSE,include=FALSE}
import matplotlib.pyplot as plt
import plotly.express as px
import plotly.offline as py
import pandas as pd
import numpy as np


### extract data 
df = pd.read_csv('countypres_2000-2016.csv')
df_1 = df[(df['year']==2008)] 
df_2 = df[(df['year']==2012)] 
df_3 = df[(df['year']==2016)]
frames = [df_1, df_2, df_3] 
df = pd.concat(frames)

df_dem = df[df['party'] == 'democrat']
df_rep = df[df['party'] == 'republican']
frames = [df_dem, df_rep] 
df = pd.concat(frames)

df['won_proportion'] = df['candidatevotes']/df['totalvotes']
df = df[['year', 'state', 'state_po', 'FIPS', 'county', 'party', 'won_proportion']]
df = df.dropna(subset=['FIPS'])
df = df.astype({'FIPS': int})
df = df.astype({'FIPS': str})
### fill with 0
df['FIPS'] = df['FIPS'].apply(lambda x: x.zfill(5))
df = df.reset_index(drop = True)
df.to_csv('final_election.csv')

df_unemploy = pd.read_excel('Unemployment.xls', sheet_name = 'Unemployment Med HH Income', skiprows = 7)
df_unemploy = df_unemploy[['FIPStxt', 'Stabr', 'area_name', 
                                       'Unemployment_rate_2008', 'Unemployment_rate_2012', 'Unemployment_rate_2016']]
df_unemploy = pd.melt(df_unemploy, id_vars=["FIPStxt", 'Stabr', 'area_name'], var_name="year", value_name="unemployment_rate")
df_unemploy = df_unemploy.rename(columns={"FIPStxt": "FIPS"})
df_unemploy.year = df_unemploy.year.apply(lambda x: x[-4:])
df_unemploy = df_unemploy.astype({'FIPS': str})
df_unemploy['FIPS'] = df_unemploy['FIPS'].apply(lambda x: x.zfill(5))
df_unemploy = df_unemploy.astype({'year': str})
df_unemploy = df_unemploy.astype({'FIPS': str})
df_unemploy.to_csv('final_unemployment.csv')

```






## Task 2

#### Note:
Three States that I choose is Idaho, Colorado, and California
I choose democrat Candidates 

We can see that as the color become darker which means the candidates have large vote percentage (notices the scale is different for different states)
Same with the unemployment rate. As the color become darker the unemployment rate grows. 

```{r,echo=FALSE,warning=FALSE,include=FALSE}
new_election_data = read_csv('final_election.csv')
new_unemployment_data = read.csv('final_unemployment.csv')
new_unemployment_data$FIPS = as.character(new_unemployment_data$FIPS)
new_unemployment_data$FIPS = str_pad(new_unemployment_data$FIPS, 5, pad = "0")
```


```{r,echo=FALSE,warning=FALSE,include=FALSE}
# load shape file 
shape = st_read('tl_2019_us_county.shp')

map_co = shape %>% filter(STATEFP =='08') %>% rename('FIPS'='GEOID')

map_ca = shape %>% filter(STATEFP =='06') %>% rename('FIPS'='GEOID')

map_id = shape %>% filter(STATEFP =='16') %>% rename('FIPS'='GEOID')
```

## Year of 2008

```{r fig1, fig.height = 30, fig.width = 25, fig.align = "center",echo=FALSE,warning=FALSE}
# CA Election Vote in 2008
election_ca_2008 = new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='CA') %>% filter(year=='2008') %>% filter(party=='democrat')
# merge with shape file 
election_ca_2008 = map_ca %>% left_join(election_ca_2008[,c('FIPS',"won_proportion")],by='FIPS') 

election_ca_2008_map = ggplot(election_ca_2008) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "magma", direction = -1)+labs(title='Democrat Vote in California in 2008', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


# CA Unemployment 
unemploy_ca_2008=new_unemployment_data %>% filter(Stabr=='CA') %>% filter(year=='2008')
unemploy_ca_2008=map_ca %>% left_join(unemploy_ca_2008, by='FIPS')
unemploy_ca_2008_map = ggplot(unemploy_ca_2008)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno",direction = -1)+labs(title='Unemployment Rate in California in 2008', fill='Unemployment Rate(%)') + 
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


# CO Election Vote
election_co_2008=new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='CO') %>% filter(year=='2008') %>% filter(party=='democrat')
election_co_2008 = map_co %>% left_join(election_co_2008[,c('FIPS',"won_proportion")],by='FIPS') 
election_co_2008_map = ggplot(election_co_2008) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "magma", direction = -1)+labs(title='Democrat Vote in Colorado in 2008', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


# CO Unemployment 
unemploy_co_2008=new_unemployment_data %>% filter(Stabr=='CO') %>% filter(year=='2008')
unemploy_co_2008=map_co %>% left_join(unemploy_co_2008, by='FIPS')
unemploy_co_2008_map = ggplot(unemploy_co_2008)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno",direction = -1)+labs(title='Unemployment Rate in Colorado in 2008', fill='Unemployment Rate(%)')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


# ID Election Vote
election_id_2008=new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='ID') %>% filter(year=='2008') %>% filter(party=='democrat')
election_id_2008 = map_id %>% left_join(election_id_2008[,c('FIPS',"won_proportion")],by='FIPS') 
election_id_2008_map = ggplot(election_id_2008) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "magma", direction = -1)+labs(title='Democrat Vote in Idaho in 2008', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())

# ID Unemployment 
unemploy_id_2008=new_unemployment_data %>% filter(Stabr=='ID') %>% filter(year=='2008')
unemploy_id_2008=map_id %>% left_join(unemploy_id_2008, by='FIPS')
unemploy_id_2008_map = ggplot(unemploy_id_2008)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno", direction = -1)+labs(title='Unemployment Rate in Idaho in 2008', fill='Unemployment Rate(%)')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


ggarrange(election_ca_2008_map, unemploy_ca_2008_map, election_co_2008_map, unemploy_co_2008_map, election_id_2008_map, unemploy_id_2008_map,
          ncol = 2, nrow=3, align = 'hv', widths = c(6,6,6,6,6,6))

```




## Year of 2012

```{r fig2, fig.height = 30, fig.width = 25, fig.align = "center",echo=FALSE,warning=FALSE}
# CA Election Vote
election_ca_2012 = new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='CA') %>% filter(year=='2012') %>% filter(party=='democrat')
election_ca_2012 = map_ca %>% left_join(election_ca_2012[,c('FIPS',"won_proportion")],by='FIPS') 
election_ca_2012_map = ggplot(election_ca_2012) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "magma", direction = -1)+labs(title='Democrat Vote in California in 2012', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


# CA Unemployment 
unemploy_ca_2012=new_unemployment_data %>% filter(Stabr=='CA') %>% filter(year=='2012')
unemploy_ca_2012=map_ca %>% left_join(unemploy_ca_2012, by='FIPS')
unemploy_ca_2012_map = ggplot(unemploy_ca_2012)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno",direction = -1)+labs(title='Unemployment Rate in California in 2012', fill='Unemployment Rate(%)') + 
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


# CO Election Vote
election_co_2012=new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='CO') %>% filter(year=='2012') %>% filter(party=='democrat')
election_co_2012 = map_co %>% left_join(election_co_2012[,c('FIPS',"won_proportion")],by='FIPS') 
election_co_2012_map = ggplot(election_co_2012) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "magma", direction = -1)+labs(title='Democrat Vote in Colorado in 2012', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


# CO Unemployment 
unemploy_co_2012=new_unemployment_data %>% filter(Stabr=='CO') %>% filter(year=='2012')
unemploy_co_2012=map_co %>% left_join(unemploy_co_2012, by='FIPS')
unemploy_co_2012_map = ggplot(unemploy_co_2012)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno",direction = -1)+labs(title='Unemployment Rate in Colorado in 2012', fill='Unemployment Rate(%)')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


# ID Election Vote
election_id_2012=new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='ID') %>% filter(year=='2012') %>% filter(party=='democrat')
election_id_2012 = map_id %>% left_join(election_id_2012[,c('FIPS',"won_proportion")],by='FIPS') 
election_id_2012_map = ggplot(election_id_2012) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "magma", direction = -1)+labs(title='Democrat Vote in Idaho in 2012', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())

# ID Unemployment 
unemploy_id_2012=new_unemployment_data %>% filter(Stabr=='ID') %>% filter(year=='2012')
unemploy_id_2012=map_id %>% left_join(unemploy_id_2012, by='FIPS')
unemploy_id_2012_map = ggplot(unemploy_id_2012)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno", direction = -1)+labs(title='Unemployment Rate in Idaho in 2012', fill='Unemployment Rate(%)')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


ggarrange(election_ca_2012_map, unemploy_ca_2012_map, election_co_2012_map, unemploy_co_2012_map, election_id_2012_map, unemploy_id_2012_map,
          ncol = 2, nrow=3, align = 'hv', widths = c(6,6,6,6,6,6))

```






## Year of 2016

```{r fig3, fig.height = 30, fig.width = 25, fig.align = "center" ,echo=FALSE,warning=FALSE}
# CA Election Vote
election_ca_2016 = new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='CA') %>% filter(year=='2016') %>% filter(party=='democrat')
election_ca_2016 = map_ca %>% left_join(election_ca_2016[,c('FIPS',"won_proportion")],by='FIPS') 
election_ca_2016_map = ggplot(election_ca_2016) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "magma", direction = -1)+labs(title='Democrat Vote in California in 2016', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


# CA Unemployment 
unemploy_ca_2016=new_unemployment_data %>% filter(Stabr=='CA') %>% filter(year=='2016')
unemploy_ca_2016=map_ca %>% left_join(unemploy_ca_2016, by='FIPS')
unemploy_ca_2016_map = ggplot(unemploy_ca_2016)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno",direction = -1)+labs(title='Unemployment Rate in California in 2016', fill='Unemployment Rate(%)') + 
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


# CO Election Vote
election_co_2016=new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='CO') %>% filter(year=='2016') %>% filter(party=='democrat')
election_co_2016 = map_co %>% left_join(election_co_2016[,c('FIPS',"won_proportion")],by='FIPS') 
election_co_2016_map = ggplot(election_co_2016) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "magma", direction = -1)+labs(title='Democrat Vote in Colorado in 2016', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


# CO Unemployment 
unemploy_co_2016=new_unemployment_data %>% filter(Stabr=='CO') %>% filter(year=='2016')
unemploy_co_2016=map_co %>% left_join(unemploy_co_2016, by='FIPS')
unemploy_co_2016_map = ggplot(unemploy_co_2016)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno",direction = -1)+labs(title='Unemployment Rate in Colorado in 2016', fill='Unemployment Rate(%)')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())

# ID Election Vote
election_id_2016=new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='ID') %>% filter(year=='2016') %>% filter(party=='democrat')
election_id_2016 = map_id %>% left_join(election_id_2016[,c('FIPS',"won_proportion")],by='FIPS') 
election_id_2016_map = ggplot(election_id_2016) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "magma", direction = -1)+labs(title='Democrat Vote in Idaho in 2016', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())
# ID Unemployment 
unemploy_id_2016=new_unemployment_data %>% filter(Stabr=='ID') %>% filter(year=='2016')
unemploy_id_2016=map_id %>% left_join(unemploy_id_2016, by='FIPS')
unemploy_id_2016_map = ggplot(unemploy_id_2016)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno", direction = -1)+labs(title='Unemployment Rate in Idaho in 2016', fill='Unemployment Rate(%)')+
  xlab('Longitude')+ylab('Latitude')+  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())


ggarrange(election_ca_2016_map, unemploy_ca_2016_map, election_co_2016_map, unemploy_co_2016_map, election_id_2016_map, unemploy_id_2016_map,
          ncol = 2, nrow=3, align = 'hv', widths = c(6,6,6,6,6,6))

```























